\documentclass[11pt,a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage{isabelle,isabellesym}

\usepackage{amssymb}
  %for \<leadsto>, \<box>, \<diamond>, \<sqsupset>, \<mho>, \<Join>,
  %\<lhd>, \<lesssim>, \<greatersim>, \<lessapprox>, \<greaterapprox>,
  %\<triangleq>, \<yen>, \<lozenge>
\usepackage{amsmath}

\usepackage[ruled]{algorithm2e}

\SetKwInput{Init}{Initialization}
\SetKwInput{Online}{Online Matching}
\SetKwFor{Arrival}{On arrival of}{}{}
\SetKwIF{If}{ElseIf}{Else}{if}{}{else if}{else}{}

% this should be the last package used
\usepackage{pdfsetup}

% urls in roman style, theory text in math-similar italics
\urlstyle{rm}
\isabellestyle{it}

\begin{document}

\title{A Formal Analysis of RANKING}
\author{Mohammad Abdulaziz and Christoph Madlener}
\maketitle

\begin{abstract}
We describe a formal correctness proof of RANKING, an online algorithm for online bipartite8
matching. An outcome of our formalisation is that it shows that there is a gap in all combinatorial9
proofs of the algorithm. Filling that gap constituted the majority of the effort which went into10
this work. This is despite the algorithm being one of the most studied algorithms and a central11
result in theoretical computer science. This gap is an example of difficulties in formalising graphical12
arguments which are ubiquitous in the theory of computing.
\end{abstract}

\tableofcontents

% sane default for proof documents
\parindent 0pt\parskip 0.5ex

\section{Introduction}
In the online bipartite matching problem we are given a bipartite graph $G=(U,V,E)$. The \emph{online}
vertices in $U$ arrive one by one, revealing the edges incident to them in $G$ on arrival. On
arrival of a vertex $u \in U$ we have to make the irrevokable decision to match it to one of
its \emph{offline} neighbors in $V$ (or not match it at all).

Karp et al.\ give a simple randomized algorithm, and prove its optimality~\cite{karp1990}. The
algorithm chooses a random \emph{ranking} (think total order) of the offline vertices $V$ and
matches any arriving $u \in U$ to the neighbor with the lowest rank (if possible):

\begin{algorithm}[H]
\DontPrintSemicolon
\caption{RANKING}\label{alg:ranking}
\Init{Choose a random permutation (ranking) $\sigma$ of $V$}
\Online{}
\Arrival{$u \in U$}{
  $N(u) \gets \text{set of unmatched neighbors of }u$\\
  \If{$N(u) \neq \emptyset$}{
    match $u$ to the vertex $v \in N(u)$ that minimizes $\sigma(v)$
  }
}
\end{algorithm}

Optimal in this case means that the algorithm reaches a competitive ratio of $1 - \frac{1}{e}$,
the best possible for online bipartite matching~\cite{karp1990}.
We give a formal proof that RANKING reaches a competitive ratio of $1-\frac{1}{e}$,
 following a proof due to Birnbaum and Mathieu~\cite{birnbaum2008}.

\hyperref[sec:prelims]{Section~\ref*{sec:prelims}} introduces necessary definitions and auxiliary lemmas for the rest of the
proof. In~\autoref{sec:ranking} we consider RANKING in a deterministic setting, i.e.\ we fix
an arbitrary ranking for the offline vertices. This section also includes the most involved part
of the formalization; proving for the first time that the competitve ratio is determined by instances where a perfect
matching exists in $G$. Finally, \autoref{sec:prob} states RANKING as a randomized algorithm and
deals with the probabilistic parts of the proof, yielding the competitive ratio in the end.

\section{Preliminaries}\label{sec:prelims}
We start by providing auxiliary material for relevant concepts used in the formalization.
In particular we need additional theory for lists, the model for permutations we 
chose~(\autoref{sec:more-list}).
In~\autoref{sec:more-graph} we define -- and prove properties of -- relevant graph theory.
Finally, in~\autoref{sec:misc} auxiliary lemmas for the probabilistic part of the formalization
are provided.

% generated text of all theories
\input{session}

% optional bibliography
\bibliographystyle{abbrv}
\bibliography{root}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
